---
title: "census_data"
author: "Dario"
date: "3/19/2021"
output: html_document
---

## Census data preparation

Source: U.S. Census Bureau, 2015-2019 American Community Survey 5-Year Estimates

Raw data shows median earnings in the past 12 months (in 2019, inflation-adjusted dollars) by sex and by educational attainment for the population 25 years and over.

```{r}
raw_data <- read.csv("ACSDT5Y2019.B20004_data_with_overlays_2021-03-18T142834.csv", skip = 1, header = TRUE, sep = ",", quote = "\"", stringsAsFactors = FALSE)
head(raw_data)
```
**Rows 16 to 38 not needed for our task -> delete**
```{r}
data <- raw_data
data[,(15:38)] <- NULL
data[,1]<-NULL
head(data)
```
**cleaning zip code**
```{r}
colnames(data)[1]  <- "zip"
data$zip <- gsub("ZCTA5 ","",data$zip)
head(data)
```

Explanation of Symbols:

  - "**" in margin of error: indicates that either no sample observations or too few sample observations were        available
  - "-" entry in the estimate: indicates that either no sample observations or too few sample observations were        available to compute an estimate, or a ratio of medians cannot be calculated because one or both of the          median estimates falls in the lowest interval or upper interval of an open-ended distribution, or the            margin of error associated with a median was larger than the median itself.
  - "-" following a median estimate means the median falls in the lowest interval of an open-ended distribution.
  - "+" following a median estimate means the median falls in the upper interval of an open-ended distribution.
  - "***" entry in the margin of error: indicates that the median falls in the lowest interval or upper interval     of an open-ended distribution.
  - "*****" entry in the margin of error column indicates that the estimate is controlled.
  - "N" entry in the estimate and margin of error columns indicates that data for this geographic area cannot be     displayed because the number of sample cases is too small.
  - "(X)" means that the estimate is not applicable or not available.


**turn symbols to NA for easier handling**
```{r}
#couldn't figure out one-liner
data[ data == "-" ] <- NA
data[ data == "**" ] <- NA
data[ data == "+" ] <- NA
data[ data == "***" ] <- NA
data[ data == "*****" ] <- NA
data[ data == "N" ] <- NA
data[ data == "X" ] <- NA
head(data)
```
```{r}
data_ziptotal= data
data_ziptotal[,4:14]<-NULL
head(data_ziptotal)
```


**data frame without NA values - omit rows containing one NA or more**
```{r}
data_lean <- data_ziptotal[complete.cases(data_ziptotal),]
head(data_lean)
```
**check how much left**
```{r}
nrow(data_ziptotal)
nrow(data_lean)
nrow(data_lean)/nrow(data_ziptotal)
```




