---
title: "Tidycensus_data_prep"
author: "Dario"
date: "3/31/2021"
output: html_document
---
## Install packages 

```{r cars}
install.packages("githubinstall")
library(githubinstall)
githubinstall("tidycensus")
install.packages(tidyverse)
```

## Load Packages

```{r}
library(tidycensus)
library(tidyverse)
```
```{r}
census_api_key("f546cab669d39b586c2d8d823eb75d9bea4c7668", install = TRUE)
```

## Look up variables

```{r}
acs5_vars <- load_variables(2019,"acs5",cache=TRUE)
View(acs5_vars)
```

B19013_001: Median houshold income past 12 months (inflation adjusted)

B20004_001 : Median earnings past 12 months (inflation adjusted)

B06011_001 : Median estimated income past 12 months (inflation adjusted)

B06009_001: place of birth by educational income (total eg. number for place of birth?) 

## Load data

```{r}
incm_zip <- get_acs(geography = "zcta", variables = "B06011_001", year = 2019)
incm_state<- get_acs(geography = "state", variables = "B06011_001", year = 2019)
head(incm_zip)
head(incm_state)
```


## Clean data in best way for merge with twitter data

  - NA values still in table
  - yet to do....

```{r}

```








